---
editor_options:
  markdown:
    wrap: 72
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Sofiia Folvarochna, Oleh Hykavyi, Nadiia Lakoma*

## Introduction

During the past three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations.

## Data description

There are 5 datasets uploaded on the cms.

To determine your variant, take your team number from the list of teams
on cms and take *mod 5* - this is the number of your data set.

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.

-   **1 - discrimination** This data set consists of tweets that have
    discriminatory (sexism or racism) messages or of tweets that are of
    neutral mood. The task is to determine whether a given tweet has
    discriminatory mood or does not.

-   **2 - fake news** This data set contains data of American news: a
    headline and an abstract of the article. Each piece of news is
    classified as fake or credible. The task is to classify the news
    from test.csv as credible or fake.

-   **3 - sentiment** All the text messages contained in this data set
    are labeled with three sentiments: positive, neutral or negative.
    The task is to classify some text message as the one of positive
    mood, negative or neutral.

-   **4 - spam** This last data set contains SMS messages classified as
    spam or non-spam (ham in the data set). The task is to determine
    whether a given message is spam or non-spam.

Each data set consists of two files: *train.csv* and *test.csv*. The
first one you will need find the probabilities distributions for each of
the features, while the second one is needed for checking how well your
classifier works.

```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
install.packages("wordcloud")
library(wordcloud)
install.packages("RColorBrewer")
library(RColorBrewer)
install.packages("wordcloud2")
library(wordcloud2)
```
## Instructions

-   The first step is data pre-processing, which includes removing
    punctuation marks and stop words

-   represent each message as a bag-of-words

-   using the training set, calculate all the conditional probabilities
    in formula (1)

-   use those to predict classes for messages in the test set

-   evaluate effectiveness of the classifier by calculating the
    corresponding metrics

-   shortly summarize your work

-   do not forget to submit both the (compiled) Rmd source file and the .html
    output
    
### Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

```{r}
list.files(getwd())
list.files("data/1-discrimination")
```

```{r}
test_path <- "data/1-discrimination/test.csv"
train_path <- "data/1-discrimination/train.csv"
stop_words <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
splitted_stop_words <- strsplit(stop_words, split = "\r\n")
splitted_stop_words <- splitted_stop_words[[1]]
stop_words <- splitted_stop_words
```


```

```{r}
train <- read.csv(file = train_path, stringsAsFactors = FALSE)
test <- read.csv(file = test_path, stringsAsFactors = FALSE)
discrim_and_neutral <- table(train$label)
prob_of_neutral <- discrim_and_neutral[2] / sum(discrim_and_neutral[1], discrim_and_neutral[2])
prob_of_discrim <- discrim_and_neutral[1] / (sum(discrim_and_neutral[1], discrim_and_neutral[2]))
```

```{r}
# note the power functional features of R bring us!
df_neutral <- train[(train$label == "neutral"), ]
df_discrim <- train[(train$label == "discrim"), ]
tidy_text_discrim <- unnest_tokens(df_discrim, "splitted", "tweet", token = "words") %>% filter(!splitted %in% stop_words)
df_tidy_text_discrim <- tidy_text_discrim %>% count(splitted, sort = TRUE)
tidy_text_neutral <- unnest_tokens(df_neutral, "splitted", "tweet", token = "words") %>% filter(!splitted %in% stop_words)
df_tidy_text_neutral <- tidy_text_neutral %>% count(splitted, sort = TRUE)
amout_of_neutral_words <- sum(df_tidy_text_neutral[, "n"])
amout_of_discrim_words <- sum(df_tidy_text_discrim[, "n"])
tidy_text <- unnest_tokens(train, "splitted", "tweet", token = "words") %>% filter(!splitted %in% stop_words)
amount_of_all_words <- nrow(tidy_text %>% count(splitted, sort = TRUE))
```

### Data visualization

Each time you work with some data, you need to understand it before you
start processing it. R has very powerful tools to make nice plots and
visualization. Show what are the most common words for negative and
positive examples as a histogram, word cloud etc. Be creative!
```{r}
set.seed(1234) # for reproducibility
wordcloud2(data = df_tidy_text_discrim[c(1:50), c(1:2)], size = 2, color = "random-dark", backgroundColor = "lightyellow")
wordcloud2(data = df_tidy_text_neutral[c(1:50), c(1:2)], size = 2, color = "random-dark", backgroundColor = "lightyellow")
```

## Classifier implementation

```{r}
naiveBayes <- setRefClass("naiveBayes",

    # here it would be wise to have some vars to store intermediate result
    # frequency dict etc. Though pay attention to bag of wards!
    fields = list(),
    methods = list(
        # prepare your training data as X - bag of words for each of your
        # messages and corresponding label for the message encoded as 0 or 1
        # (binary classification task)
        fit = function(X, y) {
            X$prob <- (X$n + 1) / (y + amount_of_all_words)
            return(X)
        },

        # return prediction for a single message
        predict = function(message) {
            # TODO
            tst <- message
            msg <- c(tst)
            df_msg <- data.frame(msg)
            df_msg <- unnest_tokens(df_msg, "word", "msg", token = "words") %>% filter(!word %in% stop_words)
            prob_of_n <- prob_of_neutral
            for (x in df_msg[, 1]) {
                if (identical(which(neutral$splitted == x), integer(0))) {
                    prob_of_n <- prob_of_n * (1 / (amout_of_neutral_words + amount_of_all_words))
                } else {
                    prob_of_n <- prob_of_n * neutral[which(neutral$splitted == x), 3]
                }
            }
            prob_of_d <- prob_of_discrim
            for (x in df_msg[, 1]) {
                if (identical(which(discrim$splitted == x), integer(0))) {
                    prob_of_d <- prob_of_d * (1 / (amout_of_discrim_words + amount_of_all_words))
                } else {
                    prob_of_d <- prob_of_d * discrim[which(discrim$splitted == x), 3]
                }
            }
            if (prob_of_n > prob_of_d) {
                return("neutral")
            } else {
                return("discrim")
            }
        },

        # score you test set so to get the understanding how well you model
        # works.
        # look at f1 score or precision and recall
        # visualize them
        # try how well your model generalizes to real world data!
        score = function() {
            count_d_true <- 0
            count_d_false <- 0
            count_n_true <- 0
            count_n_false <- 0
            # TODO
            res_list <- list()
            for (x in 1:nrow(test)) {
                predict_of_msg <- model$predict(test[x, 4])
                res_list[[x]] <- predict_of_msg
                if (predict_of_msg == test[x, 3]) {
                    if (test[x, 3] == "neutral") {
                        count_n_true <- count_n_true + 1
                    } else {
                        count_d_true <- count_d_true + 1
                    }
                } else {
                    if (test[x, 3] == "neutral") {
                        count_n_false <- count_n_false + 1
                    } else {
                        count_d_false <- count_d_false + 1
                    }
                }
            }
            slices <- c(count_n_true, count_n_false)
            lbls <- c("Right", "Wrong")
            slices1 <- c(count_d_true, count_d_false)
            lbls1 <- c("Right", "Wrong")
            jpeg("Visualisation_of_N.jpg")
            par(mfrow = c(2, 1))
            pie(slices, lbls, main = "Visualisation of mistakes in neutral messages")
            dev.off()
            jpeg("Visualisation_of_D.jpg")
            par(mfrow = c(2, 1))
            pie(slices1, lbls1, main = "Visualisation of mistakes in discriminative messages")
            dev.off()
        }
    )
)

model <- naiveBayes()
neutral <- model$fit(df_tidy_text_neutral, amout_of_neutral_words)
discrim <- model$fit(df_tidy_text_discrim, amout_of_discrim_words)
# write.csv(neutral, "data_n.csv")
# write.csv(discrim, "data_d.csv")
model$score()
```

## Measure effectiveness of your classifier
-   Note that accuracy is not always a good metric for your classifier.
    Look at precision and recall curves, F1 score metric.
-   Visualize them.
-   Show failure cases.

Visualisation is saved in to .jpg files. In the "Visualisation_of_N" we have 
ratio of right and wrong classified neutral tweets. In the second one, 
the same thing for discriminative tweets.

## Conclusions

Summarize your work by explaining in a few sentences the points listed
below.

-   Describe the method implemented in general. Show what are
    mathematical foundations you are basing your solution on.
-   List pros and cons of the method. This should include the
    limitations of your method, all the assumption you make about the
    nature of your data etc.

Our conclusion:
To perform this task, we used the product rule. This helps us find the 
probability of the occurrence of a word in each type of message 
(neutral or discriminatory) and then compare this data and return the 
type with a higher probability. This approach is good but quite primitive 
because it is guided by a relatively small database and does not study 
the nature of messages but simply counts the popularity of words.
Due to the large number of incorrectly selected types in train.csv,
our algorithm works correctly, but some messages are still decoded incorrectly.
Also, stop_words.txt should be much broader and leave only reference words 
that will facilitate the algorithm's work in type recognition.

